---
title: "Unique unrecoverable calls"
output: html_notebook
---

```{r}
library(tidyverse)
library(DBI)
library(RSQLite)
```

```{r}
conn = dbConnect(RSQLite::SQLite(), dbname='../processed_data/all_variants.sqlite')
dbExecute(conn, 'PRAGMA cache_size=-4000000')
dbExecute(conn, 'PRAGMA temp_store=MEMORY')
dbExecute(conn, 'PRAGMA query_only=1')
dbListTables(conn)
```

Read in the data
```{r}
gdc_unrecoverable_unique_tbl <- read_tsv(
    '../processed_data/gdc_not_recoverable_unique_variants.tsv.gz',
    col_types = cols(
        .default = col_character(),
        start_position = col_integer(), end_position = col_integer(),
        mc3_t_depth = col_integer(), mc3_t_ref_count = col_integer(), mc3_t_alt_count = col_integer(),
        mc3_n_depth = col_integer(), mc3_n_ref_count = col_integer(), mc3_n_alt_count = col_integer()
    )
) %>% filter(gdc_variant_type == 'SNP')

mc3_unrecoverable_unique_tbl <- read_tsv(
    '../processed_data/mc3_not_recoverable_unique_variants.tsv.gz',
    col_types = cols(
        .default = col_character(),
        start_position = col_integer(), end_position = col_integer(),
        mc3_t_depth = col_integer(), mc3_t_ref_count = col_integer(), mc3_t_alt_count = col_integer(),
        mc3_n_depth = col_integer(), mc3_n_ref_count = col_integer(), mc3_n_alt_count = col_integer()
    )
) %>% filter(mc3_variant_type == 'SNP')
```

```{r}
wga_samples <- dbGetQuery(
        conn, 
        "SELECT DISTINCT tumor_sample_barcode FROM full_overlap WHERE instr(gdc_gdc_filter, 'wga') > 0"
    ) %>% pull(tumor_sample_barcode)
```

## MC3 unrecoverable unique calls
```{r}
mc3_tbl <- mc3_unrecoverable_unique_tbl %>%
    mutate(
        mc3_by_muse = as.integer(str_detect(mc3_callers, 'MUSE')),
        mc3_by_mutect = as.integer(str_detect(mc3_callers, 'MUTECT')),
        mc3_by_somaticsniper = as.integer(str_detect(mc3_callers, 'SOMATICSNIPER')),
        mc3_by_radia = as.integer(str_detect(mc3_callers, 'RADIA')),
        mc3_by_varscan = as.integer(str_detect(mc3_callers, 'VARSCAN')),
        mc3_by_pindel = as.integer(str_detect(mc3_callers, 'PINDEL')),
        mc3_by_indelocator = as.integer(str_detect(mc3_callers, 'INDELOCATOR')),
        is_wga = ifelse(tumor_sample_barcode %in% wga_samples, 'Y', 'N')
    )
```



### Difference by Caller and WGA 
```{r}
mc3_per_caller_wga_tbl <- mc3_tbl %>% 
    count(mc3_callers, is_wga, sort=TRUE) %>%
    rename(num_unrecoverable_mut = n) %>%
    group_by(is_wga) %>%
    mutate(percentage_unrecoverable = num_unrecoverable_mut / sum(num_unrecoverable_mut) * 100) %>%
    ungroup() %>%
    full_join(
        as.tibble(dbGetQuery(
            conn, 
            "
            WITH tab AS (
                SELECT 
                    mc3_callers, mc3_ncallers,
                    CASE WHEN mc3_filter IS NULL THEN 'N' 
                         WHEN instr(mc3_filter, 'wga') > 0 THEN 'Y' 
                         ELSE 'N' END AS is_wga 
                FROM full_overlap 
                WHERE shared_by_gdc_mc3 = 1 OR only_in_mc3 = 1
            )
            SELECT mc3_callers, mc3_ncallers, is_wga, count(*) AS num_mut
            FROM tab                  
            GROUP BY mc3_callers, is_wga
            "
        )) %>%
            group_by(is_wga) %>%
            mutate(percentage = num_mut / sum(num_mut) * 100) %>%
            ungroup(),
        by = c("mc3_callers", "is_wga")
    ) %>%
    # Fill NA with 0
    mutate_all(funs(replace(., is.na(.), 0))) %>%
    select(mc3_callers, mc3_ncallers, is_wga, percentage_unrecoverable, percentage, num_unrecoverable_mut, num_mut)

mc3_per_caller_wga_tbl
```

Make sure the percentages sum up to 1
```{r}
mc3_per_caller_wga_tbl %>% group_by(is_wga) %>% summarize(sum(percentage_unrecoverable), sum(percentage))
```

Plot it
```{r, fig.width=6, fig.height=3.5}
big_percent_two_mc3_callers <- mc3_per_caller_wga_tbl %>% 
    filter(percentage >= 5 | percentage_unrecoverable >= 5, mc3_ncallers == 2) %>% 
    pull(mc3_callers)

plot_tbl <- mc3_per_caller_wga_tbl %>% 
    mutate(
        mc3_callers = factor(case_when(
            mc3_callers %in% big_percent_two_mc3_callers ~ str_replace(str_to_lower(mc3_callers), '\\|', '\n'),
            mc3_ncallers == 2 ~ 'other 2 callers',
            mc3_ncallers == 3 ~ 'any 3 callers',
            mc3_ncallers >= 4 ~ 'any ≥4 callers'
        ), levels = c('somaticsniper\nvarscans', 'radia\nvarscans', 'mutect\nmuse', 
                      'other 2 callers', 'any 3 callers', 'any ≥4 callers')
        )
    ) %>% 
    group_by(mc3_callers, is_wga) %>%
    summarize(
        percentage_unrecoverable = sum(percentage_unrecoverable), 
        percentage = sum(percentage)
    ) %>%
    rename(unrecoverable=percentage_unrecoverable, shared=percentage) %>%
    gather(key = 'type', value = 'percentage', -mc3_callers, -is_wga) %>%
    mutate(
        type = factor(
            type, 
            levels = c('unrecoverable', 'shared')
        )
    ) %>%
    arrange(percentage)

ggplot(plot_tbl, aes(x = mc3_callers, fill = type, y = percentage)) + 
    geom_col(position = 'dodge') + 
    scale_fill_manual(
        name = NULL, 
        values = c('unrecoverable' = '#ED909F', 'shared' = 'gray30'),
        labels = c('MC3 unrecoverable unique', 'All MC3 calls')
    ) + 
    facet_grid(~ is_wga, labeller = labeller(is_wga = c('N' = 'Not WGA samples', 'Y' = 'WGA samples'), scales = 'fixed_y')) + 
    scale_y_continuous(labels = scales::percent_format(scale = 1)) + 
    labs(x = NULL, y = 'Percentage of total SNV calls') + 
    theme_bw() + 
    theme(
        axis.text.x = element_text(angle = 35, hjust = 1, vjust = 1),
        legend.justification = c(0, 1),
        legend.position = c(0, 1),
        legend.background = element_blank()
    )

ggsave('../figures/MC3_unrecoverable_unique_snv_calls_per_caller.pdf', width = 6, height = 3.5)
```


### Exploring possible sources of difference
filter out WGA calls
```{r}
mc3_tbl <- mc3_tbl %>% filter(is_wga == 'N')
```

filter difference
```{r}
mc3_tbl %>% count(mc3_filter, sort = TRUE)
```

Calculate the VAF
```{r}
x_breaks <- seq(from = 0, to = 1, by =  0.05)

plot_tbl <- mc3_tbl %>% 
    mutate(
        mc3_vaf = as.integer(mc3_t_alt_count) / as.integer(mc3_t_depth)
    )

ggplot(plot_tbl, aes(x=mc3_vaf)) + 
    geom_histogram(breaks = x_breaks)
```

Caller difference
```{r}
mc3_tbl %>% count(mc3_callers, sort = TRUE)
```

Gene difference 
```{r}
mc3_tbl %>% 
    count(mc3_hugo_symbol, sort = TRUE) %>% 
    head(10)
```

```{r}
mc3_tbl %>% count(tumor_sample_barcode, sort = TRUE)
```

Manual investigation
```{r}
mc3_tbl %>% filter(mc3_t_alt_count / mc3_t_depth >= 0.1, mc3_variant_classification == 'Missense_Mutation') %>% head(10)
```



## GDC unrecoverable unique calls
```{r}
gdc_tbl <- gdc_unrecoverable_unique_tbl %>% 
    mutate(is_wga = ifelse(!is.na(gdc_gdc_filter) & str_detect(gdc_gdc_filter, 'wga'), 'Y', 'N'))
```

```{r}
gdc_tbl %>% count(gdc_callers, sort=TRUE) %>% head()
```

```{r}
gdc_tbl %>% count(is_wga, sort = TRUE)
```


### Difference by caller and WGA 
Make sure there is a significant increase of mutect only call in WGA samples.
```{r}
sort_gdc_callers <- function(gdc_callers, ...) {
    gdc_callers_fixed_order <- c('muse', 'mutect', 'somaticsniper', 'varscan')
    gdc_by_muse <- as.integer(str_detect(gdc_callers, 'muse'))
    gdc_by_mutect <- as.integer(str_detect(gdc_callers, 'mutect'))
    gdc_by_somaticsniper <- as.integer(str_detect(gdc_callers, 'somaticsniper'))
    gdc_by_varscan <- as.integer(str_detect(gdc_callers, 'varscan'))
    str_c(gdc_callers_fixed_order[which(c(gdc_by_muse, gdc_by_mutect, gdc_by_somaticsniper, gdc_by_varscan) == 1)], collapse = '|')
}

gdc_per_caller_wga_tbl <- gdc_tbl %>% 
    count(gdc_callers, is_wga, sort=TRUE) %>%
    rename(num_unrecoverable_mut = n) %>%
    group_by(is_wga) %>%
    mutate(percentage_unrecoverable = num_unrecoverable_mut / sum(num_unrecoverable_mut) * 100) %>%
    ungroup() %>%
    full_join(
        as.tibble(dbGetQuery(
            conn, 
            "
            WITH shared AS (
                SELECT 
                    gdc_callers,
                    CASE WHEN gdc_gdc_filter IS NULL THEN 'N' 
                         WHEN instr(gdc_gdc_filter, 'wga') > 0 THEN 'Y' 
                         ELSE 'N' END AS is_wga 
                FROM full_overlap 
                WHERE shared_by_gdc_mc3 = 1 OR only_in_gdc = 1
            )
            SELECT gdc_callers, is_wga, count(*) AS num_mut
            FROM shared                    
            GROUP BY gdc_callers, is_wga
            "
        )) %>%
            group_by(is_wga) %>%
            mutate(percentage = num_mut / sum(num_mut) * 100) %>%
            ungroup(),
        by = c("gdc_callers", "is_wga")
    ) %>%
    # Fill NA with 0
    mutate_all(funs(replace(., is.na(.), 0))) %>%
    # GDC callers can have multiple orders
    mutate(gdc_callers = pmap_chr(., .f = sort_gdc_callers)) %>%
    group_by(gdc_callers, is_wga) %>%
    summarize(
        percentage_unrecoverable = sum(percentage_unrecoverable), percentage = sum(percentage),
        num_unrecoverable_mut = sum(num_unrecoverable_mut), num_mut = sum(num_mut)
    ) %>%
    arrange(-percentage_unrecoverable) %>%
    ungroup() %>%
    select(gdc_callers, is_wga, percentage_unrecoverable, percentage, num_unrecoverable_mut, num_mut)
gdc_per_caller_wga_tbl
```

Make sure the percentages sum up to 1
```{r}
gdc_per_caller_wga_tbl %>% group_by(is_wga) %>% summarize(sum(percentage_unrecoverable), sum(percentage))
```


Plot it
```{r, fig.width=6, fig.height=3.5}
plot_tbl <- gdc_per_caller_wga_tbl %>%
    mutate(
        # Break down the caller combinations
        gdc_callers = factor(
            case_when(
                gdc_callers == 'muse|mutect|somaticsniper|varscan' ~ 'all 4 callers',
                gdc_callers %in% c('mutect', 'varscan', 'somaticsniper', 'muse') ~ gdc_callers,
                TRUE ~ 'any other caller\ncombinations'
            ),
            levels = c('mutect', 'varscan', 'somaticsniper', 'muse', 'all 4 callers', 'any other caller\ncombinations')
        )
    ) %>%
    select(gdc_callers, is_wga, unrecoverable=percentage_unrecoverable, shared=percentage) %>%
    gather(key = 'type', value = 'percentage', -gdc_callers, -is_wga) %>%
    mutate(
        type = factor(
            type, 
            levels = c('unrecoverable', 'shared')
        )
    )

ggplot(plot_tbl, aes(x = gdc_callers, fill = type, y = percentage)) + 
    geom_col(position = 'dodge') + 
    scale_fill_manual(
        name = NULL, 
        values = c('unrecoverable' = '#80A9BF', 'shared' = 'gray30'),
        labels = c('GDC unrecoverable unique', 'All GDC calls')
    ) + 
    facet_grid(~ is_wga, labeller = labeller(is_wga = c('N' = 'Not WGA samples', 'Y' = 'WGA samples'))) + 
    scale_y_continuous(labels = scales::percent_format(scale = 1)) + 
    labs(x = NULL, y = 'Percentage of total SNV calls') + 
    theme_bw() + 
    theme(
        axis.text.x = element_text(angle = 35, hjust = 1),
        legend.justification = c(0, 1),
        legend.position = c(0, 1),
        legend.background = element_blank()
    )

ggsave('../figures/GDC_unrecoverable_unique_snv_calls_per_caller.pdf', width = 6, height = 3.5)
```

See if there is any special filters about those mutations:
```{r}
gdc_tbl %>% count(gdc_filter, sort = TRUE) %>% head()
```

Gene difference:
```{r}
gdc_tbl %>% count(gdc_hugo_symbol, sort = TRUE) %>% head(10)
```

